Joshua 2.0: A Toolkit for Parsing-Based Machine Translation \\with Syntax, Semirings, Discriminative Training and Other Goodies

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited Juri's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1642 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

\section{Support for Syntax-Augmented MT}
The syntax-augmented approach to hierachical phrase-based machine
translation introduced by \newcite{samt2006} aims at improving
translation quality by introducing syntactic information into the
decoding process. To this end, one side of the bitext is parsed and a
syntax-augmented grammar containing additional nonterminal symbols is
extracted. These new symbols are obtained by annotating the
nonterminals in Hiero-style rules
\begin{center}
\begin{tabular}{c}
  {\small \tt [X]} $\Rightarrow$ {\small \tt [X,1]} sans {\small \tt [X,2]} 
  $\vert$ {\tt \small [X,1]} without {\small \tt [X,2]}
\end{tabular}
\end{center}
extracted from the bitext, the nonterminals are labeled according to
which constituents cover the nonterminal span on the parsed side of
the bitext.
\begin{center}
\begin{tabular}{c}
  {\small \tt [VP]} $\Rightarrow$ {\small \tt [VBN,1]} sans {\small
    \tt [NP,2]} $\vert$ $\mathit{VBN}_1$ without {\small \tt [NP,2]} \\
  {\small \tt [NP]} $\Rightarrow$ {\small \tt [NP,1]} sans {\small
    \tt [NP,2]} 
  $\vert$ {\small \tt [NP,1]} without {\small \tt [NP,2]} 
\end{tabular}
\end{center}
To ensure good coverage, the SAMT approach concatenates syntactic
constituents or uses CCG-style labels when there is no single
constituent spanning a nonterminal. The resulting grammar

\joshua's decoder supports SAMT-style grammars as well as non-syntactic
nonterminal annotations. An SAMT grammar

>>>>

The initial release of \joshua supported only Hiero-style SCFGs, which use a single nonterminal symbol X.  This release includes support for arbitrary SCFGs, including ones that use a rich set of linguistic nonterminal symbols.  
In particular we have added support for \newcite{samt2006}'s syntax-augmented machine
translation. SAMT grammar extraction is identical to Hiero grammar extraction, except that one side of the parallel is parsed, and syntactic labels replace the X
nonterminals in Hiero-style rules.  Instead of extracting this Hiero rule from the bitext

\noindent
\begin{tabular}{c}
  {\small \tt [X]} $\Rightarrow$ {\small \tt [X,1]} sans {\small \tt [X,2]} 
  $\vert$ {\tt \small [X,1]} without {\small \tt [X,2]}
\end{tabular}

\noindent
the nonterminals can be labeled according to
which constituents cover the nonterminal span on the parsed side of
the bitext.  This constrains what types of phrases the decoder can use when producing a translation.

\noindent
\begin{tabular}{c}
  {\small \tt [VP]} $\Rightarrow$ {\small \tt [VBN]} sans {\small
    \tt [NP]} $\vert$ {\small \tt [VBN]} without {\small \tt [NP]} \\
  {\small \tt [NP]} $\Rightarrow$ {\small \tt [NP]} sans {\small
    \tt [NP]} 
  $\vert$ {\small \tt [NP]} without {\small \tt [NP]} 
\end{tabular}

\noindent
Unlike GHKM, SAMT has the same coverage as Hiero, because it allows non-constituent phrases to get syntactic
labels using CCG-style slash notation. 
The derivations created using syntactically motivated grammars exhibit more coherent syntactic structure than Hiero and typically result in better reordering, especially for languages with word orders that diverge from English, like Urdu.
